Appendix A: Stochastic simulations of noisy phenomena

### 2017-10-27 appendixA/appendixA.Rmd
use gridExtra not cowplot

sim_large <- adaptivetau::ssa.exact(init.values = c(n = 500),
                       transitions, rateF,
                       params = list(c=c, e=e, N=1000),
                       tf=30) %>% as_tibble()

theory <- 
list(small = 100,large = 1000) %>%
  map_dfr(function(N){
    data.frame(mean = N * (1-e/c)) %>%
    mutate(plus_sd = mean + sqrt(N * e / c),
           minus_sd = mean - sqrt(N * e / c))
  }, .id = "system_size")

gillespie <- bind_rows("large"=sim_large, 
                       "small"=sim_small,  
                       .id = "system_size") %>% bind_rows(theory)
write_csv(gillespie, "gillespie.csv")

read_csv("gillespie.csv", col_types = "cdiddd")  %>% 
  ggplot(aes(x = time)) + 
  geom_hline(aes(yintercept = mean), lty=2) + 
  geom_hline(aes(yintercept = minus_sd), lty=2) + 
  geom_hline(aes(yintercept = plus_sd), lty=2) + 
  geom_line(aes(y = n), col="red") +
  facet_wrap(~system_size, scales = "free_y") + 
  labs(caption = "theoretical predictions for mean and plus/minus one standard \n deviation shown in horizontal black dashed lines.")
```

@@

```{r}
df %>% 
  filter(x > 0) %>% 
  group_by(t) %>%
  summarise(ave = mean(x), sd = sd(x)) %>%
  mutate(equib = equib) %>%
  write_csv("inflation.csv")

```{r}
read_csv("inflation.csv") %>% 
  gather(statistic, value, -t) %>%
  ggplot(aes(t,value, col=statistic, lty=statistic)) + 
 # ggplot(aes(t,ave)) +
  geom_line(lwd=1) + 
  scale_color_ptol() + 
  coord_cartesian(xlim=c(100,500), ylim=c(1.2, .5))

```{r}
p <- list(r = .5, K = 2.1, Q = 5, H = .4, sigma = .04, a = 0.27, N = 1e4)

f <- function(x) x * p$r * (1 - x / p$K)
g <- function(x) p$a * x ^ p$Q / (x^p$Q + p$H^p$Q)

x <- seq(0,2, length.out = 100)
df <- data.frame(x, f = f(x), g = g(x))

ggplot(df, aes(x)) + geom_line(aes(y = f)) + geom_line(aes(y = g), col="red")

ggplot(df, aes(x)) + geom_line(aes(y = f-g)) + 
  geom_hline(aes(yintercept=0), lty=2) +
  coord_cartesian(xlim = c(0,1.5), ylim=c(-.1,.1))  
```

@@

library(cowplot)
ggdraw() +
  draw_plot(p2, x = 0, y = 0, width = 1, height = 0.5) +
  draw_plot(p1, x = 0, y = .4, width = 1, height = 0.5)

>>>>

sim_large <- adaptivetau::ssa.exact(init.values = c(n = 500),
                       transitions, rateF,
                       params = list(c=c, e=e, N=1000),
                       tf=30) %>% as_tibble()
```

```{r}
theory <- 
list(small = 100,large = 1000) %>%
  map_dfr(function(N){
    data.frame(mean = N * (1-e/c)) %>%
    mutate(plus_sd = mean + sqrt(N * e / c),
           minus_sd = mean - sqrt(N * e / c))
  }, .id = "system_size")
```

```{r}
## Bind together both simulation results and theory results
gillespie <- bind_rows("large"=sim_large, 
                       "small"=sim_small,  
                       .id = "system_size") %>% bind_rows(theory)
write_csv(gillespie, "gillespie.csv")
```

```{r}
read_csv("gillespie.csv", col_types = "cdiddd")  %>% 
  ggplot(aes(x = time)) + 
  geom_hline(aes(yintercept = mean), lty=2) + 
  geom_hline(aes(yintercept = minus_sd), lty=2) + 
  geom_hline(aes(yintercept = plus_sd), lty=2) + 
  geom_line(aes(y = n), col="red") +
  facet_wrap(~system_size, scales = "free_y") + 
  labs(caption = "theoretical predictions for mean and plus/minus one standard \n deviation shown in horizontal black dashed lines.")
```

@@

```{r}
df %>% 
  filter(x > 0) %>% 
  group_by(t) %>%
  summarise(ave = mean(x), sd = sd(x)) %>%
  mutate(equib = equib) %>%
  write_csv("inflation.csv")
```

```{r}
read_csv("inflation.csv") %>% 
  gather(statistic, value, -t) %>%
  ggplot(aes(t,value, col=statistic, lty=statistic)) + 
 # ggplot(aes(t,ave)) +
  geom_line(lwd=1) + 
  scale_color_ptol() + 
  coord_cartesian(xlim=c(100,500), ylim=c(1.2, .5))
```

@@

gridExtra::grid.arrange(p1,p2, layout_matrix = rbind(c(1,1), c(2,2)))
